\documentclass[12pt]{article}

\usepackage{caption} %interesting things with captions
\usepackage[margin=1.5in]{geometry} %mostly set margin of the report
\usepackage{tabulary} %text wrapping in tables
\usepackage{subcaption} %let me put multiple images into a single caption
\usepackage{textcomp} %let me use \textrangle{} to get '>'
\usepackage{hyperref} %links in ToC - should be mostly the last thing imported?
\hypersetup{
	urlcolor=blue
}

%images being embeded
\usepackage{graphicx}
\graphicspath{ {images/} }

%code highlighting
\usepackage{minted}
\setminted[c++]{frame=single,linenos=true,autogobble=true,numbersep=4pt,tabsize=4}
\setminted[bash]{frame=single,linenos=true,autogobble=true,numbersep=4pt,tabsize=4}
\setminted[xml]{frame=single,linenos=true,autogobble=true,numbersep=4pt,tabsize=4}

%fix a quote mark issue
\usepackage [english]{babel}
\usepackage [autostyle, english = american]{csquotes}
\MakeOuterQuote{"}


\begin{document}
\pagenumbering{gobble}
\begin{titlepage}
	\centering
	{\Huge Itterative Development\par}
	\vspace{0.25in}
	{\Large Project 1\par}
	{\scriptsize I didn't forget to change the title this time\par}
	\vspace{2in}
	{Alex Harper\par}
	\newpage
\end{titlepage}
\pagenumbering{roman}
\tableofcontents
\newpage

\section{Introduction}

I spent a long time reading the book, it has many many pages to cover for this first project.
While I was reading, I kept putting it in terms of an idea I had been wanting to do for a little while now; making a "Reddit Bot".
The idea is that you have a piece of software from time to time do automated tasks interacting with the site \href{http://reddit.com}{reddit}.
This is slightly nebulas as there are many things you could get it to do.
Really I wanted to make a fairly general wrapper around the website actions that I can make do more interesting things later.

For this paper, it is probably going to be a bit shorter than the 20 pages I normally aim for as the minimum.
The report will be about what I did and my thoughts about things, while there will be two other documents beside this report.
Since this class is supposed to be about project managment with an iterative style, I believe creating the apporpriate documents for each time section is critical for understanding.

\section{Inception}

"This section is to be short and determine only if the project should continue and what are the surface level constraints."

That is not a direct quote but in general sums up what the book says.
As such, in the document for it, many things are simple and fairly general.
While I had done some basic looking around before writing it, I did not know exactly how things were going to happen.

\subsection{Use Cases}

It differs from the examples in the book in this section because the program is supposed to run without any user interaction.
This does essentailly mean there is no "user" in our use cases, but is still in the spirit of what should be described.
As such the section is used to describe the functional parts of the program in general.

This section has what I believe to be the core ideas of what the bot needs to be able to do.
There are 4 main parts with the focus on the program being more of a framework.
While there will be a main action that it does, I am considering it more of a test use rather than the end goal.
These main actions are high level and should be broken down into more specific things during elaboration.

A 5th case is listed as a substep because it is so important; Authenticating with the servers.
The two cases for dealing directly with reddit will both directly depend on the same version of this case, and so I think it is appropriate to pull it out on it's own.

\subsection{Supplementary Specs}

These are the non-functional requirements.
Mostly these guide the behavior of the program, but are not directly what the program does.
At some point, code to address some aspects of it will need to be implemented such as limiting the rate the API gets used.
For example, limiting the rate of requests to abide by the rules can be enforced with an internal mechanism.

\subsection{Proof of Concept}

To make sure that what I was wanting to do was even possible, two small programs were made.
I honestly was not expecting to be able to use the Qt framework for this, but apparently it works well enough.
The main question it answered was how easy would it be to customize the fields for the HTTP request.
Besides a little funny buisness with first time authentication, it is simple and easy.
The second program was to see how nice the JSON objects were that Qt provided.
They work well enough that I don't feel like putting the effort into parsing things myself.

Honestly I forgot to save this code.
I am so used to having these tiny pieces be completly disposable that I didn't move them from my temp folder before reboting.
It is no real big deal for the real project because I learned what needed to already and could do better the second time around.

\section{Iteration 1}

Now we get into the guts of the program; flushing out the most important bits of the documenation and trying to get the basics of the program working.

During this part I was able to accomplish the basics of accessing reddit.
The program currently can authenticate as a user, get information using the user credentials, and restricts the API usage per time period to prevent abuse.
This is mostly what was planned to be done for this iteration with an extra bit of enforcing the rules.

\subsection{Use Cases}

I filled out the main points of two use cases as the code directly does these two.
The cases show the different thoughts and steps that should be taken (in general form) to do the right thing.

I had to take a slightly different approach than the book due to being a single person.
Normally the head people will be making the main documentation for different parts and everyone else will be getting code done.
This means that a few people grok what is supposed to happen, put it on paper, and then eventually it gets implemented.
This leads to the documentation leading ahead of the code that is implemented.
As a single person, I have to grok it before I can program it, but I might as well program it as I try to grok it to make sure I am doing things right.
This has lead to me having the documentation matching the amount of what I have programmed.

\subsection{Domain Diagram}

This encompases the whole of what interacting with the reddit website will be like (probably).
The diagram shows the replationships of different types of responses we can get from the server.
The distinctions were layed out in the reddit documentation and has each returned item is tagged with a field to let us know what is what.

For the next iteration, I will make a second domain diagram for the imgur website and possibly a third for the specific use of the robot that is being used as the test.

\subsection{Sequence Diagram}

There is only this single sequence diagram because all the others would look almost identical at this point.
(make a session, pass to the worker object, ask object to get something, get the response)
There will be more things that can get a sequence diagram for the next iteration as things will get more varied.

\section{Where am I?}

At this point, the basic work of connecting to reddit and getting responses for some basic aspects of the site is done.
\begin{itemize}
	\item The responses from the website are easily parsed by the JSON objects provided by Qt.
	\item The requests to the servers is transparently forced to follow the rules of the site using the session object.
	\item The session object at a later point could be extended to support multiple different users.
	\item Conveniance objects are ready to be filled out for the common uses, such as getting links from a subreddit.
\end{itemize}

So for the next iteration, the short list of what I want to get done.
\begin{itemize}
	\item Fill out the conveniance objects.
	\item Get a similar system setup for posting and parsing the imgur website. (likely will be a single object)
	\item Parse reddit messages and respond to them.
\end{itemize}

\newpage
\appendix
\section{redditapi.h}
\begin{minted}{c++}
#ifndef REDDITAPI_H
#define REDDITAPI_H

#include <QObject>
#include <QDir>
#include <QString>
#include <QCoreApplication>
#include <stdlib.h>
#include <stdio.h>
#include <string>
#include <iostream>
#include <QElapsedTimer>
#include <QTimer>

#include <QtNetwork/QNetworkAccessManager>
#include <QtNetwork/QNetworkRequest>
#include <QtNetwork/QNetworkReply>
#include <QUrl>
#include <QAuthenticator>
#include <QNetworkCookieJar>

#include <QJsonDocument>
#include <QJsonArray>
#include <QJsonValue>
#include <QJsonObject>

#include "configFileParser/glopConfig.h"
/*t1*/ class Reddit_Comment;
/*t2*/ class Reddit_Account;
/*t3*/ class Reddit_Link;
/*t4*/ class Reddit_Message;
/*t5*/ class Reddit_Subreddit;
/*t6*/ class Reddit_Award;
/*Listing*/ //has another member "data" who will have another object children with a row of number entries for the list indicies

class RedditSession : public QObject{
	Q_OBJECT
	static const QString clientID;
	static const QString UserAgent;
	static const QString redirectUri;
	QString API_Token; //permanent token to get a temperary token with
	QString API_Session; //temperary session key
	static QNetworkAccessManager *networkManager;

	QElapsedTimer Session_Time_Used; // to compare to for knowing when to refresh
	QElapsedTimer RequestLimit_TimeElapsed; //elapsed time since we were told anything
	int RequestLimit_RequestsLeft; // how many times the API can be used before we are abusing
	int RequestLimit_TimeToReset; // time we needed to wait for the limit to be reset

	void commonRequestSetup(QNetworkRequest &req,QString URL);
	void commonReplyWait(QNetworkReply *reply);
	void waitForRequestLimitReset();
public slots:
	// These are the only functions you should need to call
	bool parseSettings(bool allow_example_firstRun = true); // returns false if unable to get API token, probably need to run the firstTime code
	void saveSettings();
	QNetworkReply* makePOSTrequest(QString URL, QString postData);
	QNetworkReply* makeGETrequest(QString URL);
public slots:
	void firstTimeRun_example(); // prints things out and has the user do things on the command line
	QString firstTimeRun_authorizeURL(QString wanted_scopes); //have the user visit this link
	void firstTimeRun_redirectedURL(QString URL); // after the user says yes, they are sent to a new URL with info we need here
	QString scopesDescriptions();

	void testFunction(); // for me to test things with and random junk in it

	void refreshSession();
	void sendAuthentication(QNetworkReply * reply, QAuthenticator * authenticator); //should never need to be called by hand

public:
};

#endif // REDDITAPI_H
\end{minted}

\clearpage
\section{redditapi.cpp}
\begin{minted}{c++}
#include "redditapi.h"

const QString RedditSession::clientID = "f3dZ8gIetXY69Q";
const QString RedditSession::UserAgent = "pc:glop-bot:v0.0 (by /u/glop102)";
const QString RedditSession::redirectUri = "http://www.reddit.com";
//QString RedditSession::API_Token=""; //identifies that the user authorised our program
//QString RedditSession::API_Session=""; //temperary session key
//QTimer *RedditSession::Session_Time_Used = new QTimer();
QNetworkAccessManager *RedditSession::networkManager = NULL;

bool RedditSession::parseSettings(bool allow_example_firstRun){
	GlopConfig::Settings set = GlopConfig::ParseFile((QDir::homePath()+"/.redditSettings.conf") .toStdString() );
	Session_Time_Used.start();
	RequestLimit_TimeElapsed.start();
	RequestLimit_RequestsLeft = 60; // mediocre default that will be fixed on the first API use
	RequestLimit_TimeToReset = 10;

	API_Token = QString::fromStdString( set.getValueAsString("token","") );
	if(API_Token == ""){
		if(allow_example_firstRun)
			firstTimeRun_example();
		else
			printf("No Token Found - You need to do the first run code\n");
	}else{
		refreshSession();
	}
	//printf("Session token : %s\n",API_Session.toStdString().c_str());

	//testFunction();
	if(API_Session=="")
		return false;
	else
		return true;
}

void RedditSession::saveSettings(){
	GlopConfig::Settings set;
	set.values["token"] = API_Token.toStdString();
	GlopConfig::SaveToFile((QDir::homePath()+"/.redditSettings.conf") .toStdString() , set);
}

void RedditSession::firstTimeRun_example(){
	//this is an intereactive thing that runs when the settings don't seem to be all there
	// https://github.com/reddit/reddit/wiki/OAuth2
	std::string URL = firstTimeRun_authorizeURL("submit,read,history,edit,identity,privatemessages").toStdString();

	printf("This is a first time setup\n");
	printf("You need to visit the URL below and then authorize this program\n");
	printf("%s\n",URL.c_str());
	printf("\nAfter you authorize the program, it will send you to the reddit homepage\n");
	printf("\nCopy and paste the URL of the homepage to here\n");
	std::getline(std::cin,URL);
	printf("retrieving access token...\n");

	firstTimeRun_redirectedURL(QString(URL.c_str()));
}
QString RedditSession::firstTimeRun_authorizeURL(QString wanted_scopes){
	QString responseType = "code";
	QString state = "should_be_random";
	QString duration = "permanent";
	//QString wanted_scopes = "submit,read,history,edit,identity,privatemessages";

	QString URL = QString("https://www.reddit.com/api/v1/authorize?")
				  +"client_id="+clientID
				  +"&response_type="+responseType
				  +"&state="+state
				  +"&redirect_uri="+redirectUri
				  +"&duration="+duration
				  +"&scope="+wanted_scopes;
	return URL;

}
void RedditSession::firstTimeRun_redirectedURL(QString URL){
	if(URL.indexOf("error")>=0 || URL.indexOf("code=")==-1){
		printf("There was some sort of issue with the URL, try again\n");
		exit(1);
	}
	QString code;
	code = URL.right(URL.size() - URL.indexOf("code=") - QString("code=").size() ); // clean everything left of the token
	code = code.left(code.indexOf("&")); // clean everything to the right of the token
	printf("one-time code : %s\n",code.toStdString().c_str());


	QNetworkReply *reply =
	makePOSTrequest("https://www.reddit.com/api/v1/access_token",
					QString(("grant_type=authorization_code&code="+code.toStdString()+"&redirect_uri="+redirectUri.toStdString()).c_str())
					);

	//temp = QString::fromUtf8(reply->readAll());
	//printf("reply : \n%s\n",temp.toStdString().c_str());
	//QJsonDocument top = QJsonDocument::fromJson(temp.toUtf8());
	QJsonDocument top = QJsonDocument::fromJson(reply->readAll());
	if(top.isObject()){
		QJsonObject obj = top.object();
		API_Session = obj["access_token"].toString();
		API_Token = obj["refresh_token"].toString();
		saveSettings();
	}else{
		printf("Error: response from server was not valid json\n");
		exit(0);
	}
}

QNetworkReply* RedditSession::makePOSTrequest(QString URL, QString postData){
	QNetworkRequest req;
	commonRequestSetup(req,URL);
	req.setHeader(QNetworkRequest::ContentTypeHeader,"application/x-www-form-urlencoded");
	QNetworkReply *reply = networkManager->post(req,postData.toUtf8());
	commonReplyWait(reply);
	return reply;
}

QNetworkReply* RedditSession::makeGETrequest(QString URL){
	QNetworkRequest req;
	commonRequestSetup(req,URL);
	QNetworkReply *reply = networkManager->get(req);
	commonReplyWait(reply);
	return reply;
}
void RedditSession::commonRequestSetup(QNetworkRequest &req,QString URL){
	if(networkManager == NULL){
		networkManager = new QNetworkAccessManager(this);
		connect( networkManager,SIGNAL(authenticationRequired(QNetworkReply*,QAuthenticator*)) ,this,SLOT(sendAuthentication(QNetworkReply*,QAuthenticator*)) );
	}
	if(Session_Time_Used.elapsed() > 3570*1000) //in milliseconds -  30 seconds less than every hour, just to make sure we are always valid
		refreshSession();
	if(RequestLimit_RequestsLeft <= 0)
		waitForRequestLimitReset();
	req.setUrl(QUrl(URL));
	req.setHeader(QNetworkRequest::UserAgentHeader,UserAgent);
	if(! (API_Session == ""))
		//req.setRawHeader("Authorization", ("bearer "+API_Session).toUtf8());
		req.setRawHeader("Authorization", "bearer "+API_Session.toLocal8Bit());
}
void RedditSession::commonReplyWait(QNetworkReply *reply){
	QEventLoop loop;
	connect(reply, SIGNAL(finished()), &loop, SLOT(quit()));
	loop.exec();

	if(reply->hasRawHeader("x-ratelimit-remaining")){
		RequestLimit_RequestsLeft = (int)reply->rawHeader("x-ratelimit-remaining").toFloat(); // oddly toInt hates when there is a period
		RequestLimit_TimeToReset = reply->rawHeader("x-ratelimit-reset").toInt();
		RequestLimit_TimeElapsed.restart();
		//printf("found header\n");
		//printf("left %s\n",reply->rawHeader("x-ratelimit-remaining").data());
		//printf("time %s\n",reply->rawHeader("x-ratelimit-reset").data());
		//printf("left %d time %d\n",RequestLimit_RequestsLeft,RequestLimit_TimeToReset);
	}
}
void RedditSession::waitForRequestLimitReset(){
	printf("Exahsted requests for the time period, now waiting for %d seconds\n",RequestLimit_TimeToReset);
	int delta = (RequestLimit_TimeToReset*1000) - RequestLimit_TimeElapsed.elapsed();
	QEventLoop loop;
	QTimer timeLeft;
	connect(&timeLeft, SIGNAL(timeout()), &loop, SLOT(quit()));
	timeLeft.start(delta);
	loop.exec();
}

void RedditSession::sendAuthentication(QNetworkReply *reply, QAuthenticator *authenticator){
	if(! (API_Session == "")){ // should never get called - is hard set in the commonRequestSetup
		authenticator->setUser("bearer");
		authenticator->setPassword(API_Session);
		printf("authenticated with session token\n");
	}else{ // only should happen on first API call while we are trying to get the session token
		authenticator->setUser(clientID);
		authenticator->setPassword("");
		printf("authenticated with client ID\n");
	}
}

void RedditSession::refreshSession(){
	printf("Getting new session token\n");
	QString req = "grant_type=refresh_token&refresh_token=" + API_Token;
	QNetworkReply* reply = makePOSTrequest("https://www.reddit.com/api/v1/access_token",req);

	QJsonDocument top = QJsonDocument::fromJson(reply->readAll());
	if(!top.isObject()){
		printf("Error: response from server was not valid json, unable to refresh session token\n");
		exit(0);
	}
	QJsonObject obj = top.object();
	API_Session = obj["access_token"].toString();

	Session_Time_Used.restart();
	//printf("\n%s\n\n",top.toJson().data());
}

void RedditSession::testFunction(){
	//example json from the server when first authenticating the program to the user
	//{"access_token": "VnQBpiuRpH_otGfEqD30j7g_hZQ", "token_type": "bearer", "expires_in": 3600, "refresh_token": "65960967426-zaRlzkPzLgxbAf-0Vjw24O1H9sA", "scope": "edit history read submit"}
//	QByteArray input = "{\"access_token\": \"VnQBpiuRpH_otGfEqD30j7g_hZQ\", \"token_type\": \"bearer\", \"expires_in\": 3600, \"refresh_token\": \"65960967426-zaRlzkPzLgxbAf-0Vjw24O1H9sA\", \"scope\": \"edit history read submit\"}";
//	QJsonDocument top = QJsonDocument::fromJson(input);
//	if(top.isObject());
//		printf("is object\n");
//	QJsonObject obj=top.object();
//	printf("Session %s\n",obj["access_token"].toString().toStdString().c_str());
//	printf("Refresh %s\n",obj["refresh_token"].toString().toStdString().c_str());

	//------------------------------------------------------------------------------
	//a simple test for if the session token we get allows us to access things correctly
	//auto reply = makeGETrequest("https://oauth.reddit.com/api/v1/me");
	//QString temp = QString::fromUtf8(reply->readAll());
	//printf("person: %s\n",temp.toLatin1().data());

	//------------------------------------------------------------------------------
	// playing with reading data from a subreddit and what it returns
	//auto reply = makeGETrequest("https://oauth.reddit.com/r/bottest");
	//QString temp = QString::fromUtf8(reply->readAll());
	//printf("subreddit data:\n%s\n",temp.toLatin1().data());

	//printf("headers:\n");
	//for(auto x : reply->rawHeaderPairs()){
	//	printf("\t%s : %s\n",x.first.data(),x.second.data());
	//}

	//------------------------------------------------------------------------------
	//abuse test to test how well the "wait for request limit reset" functionality works to prevent accidental abuse
	//while(true){
	//	makeGETrequest("https://oauth.reddit.com/r/bottest");
	//	printf("%d reqs in %d seconds\n",RequestLimit_RequestsLeft,RequestLimit_TimeToReset);
	//}
}

QString RedditSession::scopesDescriptions(){
	return "\
creddits:         Spend my reddit gold creddits on giving gold to other users.\
modcontributors:  Add/remove users to approved submitter lists and ban/unban or mute/unmute users from subreddits I moderate.\
modmail:          Access and manage modmail via mod.reddit.com.\
modconfig:        Manage the configuration, sidebar, and CSS of subreddits I moderate.\
subscribe:        Manage my subreddit subscriptions. Manage 'friends' - users whose content I follow.\
structuredstyles: Edit structured styles for a subreddit I moderate.\
vote:             Submit and change my votes on comments and submissions.\
wikiedit:         Edit wiki pages on my behalf\
mysubreddits:     Access the list of subreddits I moderate, contribute to, and subscribe to.\
submit:           Submit links and comments from my account.\
modlog:           Access the moderation log in subreddits I moderate.\
modposts:         Approve, remove, mark nsfw, and distinguish content in subreddits I moderate.\
modflair:         Manage and assign flair in subreddits I moderate.\
save:             Save and unsave comments and submissions.\
modothers:        Invite or remove other moderators from subreddits I moderate.\
read:             Access posts and comments through my account.\
privatemessages:  Access my inbox and send private messages to other users.\
report:           Report content for rules violations. Hide & show individual submissions.\
identity:         Access my reddit username and signup date.\
livemanage:       Manage settings and contributors of live threads I contribute to.\
account:          Update preferences and related account information. Will not have access to your email or password.\
modtraffic:       Access traffic stats in subreddits I moderate.\
wikiread:         Read wiki pages through my account\
edit:             Edit and delete my comments and submissions.\
modwiki:          Change editors and visibility of wiki pages in subreddits I moderate.\
modself:          Accept invitations to moderate a subreddit. Remove myself as a moderator or contributor of subreddits I moderate or contribute to.\
history:          Access my voting history and comments or submissions I've saved or hidden.\
flair:            Select my subreddit flair. Change link flair on my submissions.";
}

\end{minted}

\clearpage
\section{main.cpp}
\begin{minted}{c++}
#include <QCoreApplication>
#include "redditapi.h"

int main(int argc, char *argv[])
{
	QCoreApplication a(argc, argv);
	RedditSession *ra = new RedditSession;
	QMetaObject::invokeMethod(ra,"parseSettings",Qt::QueuedConnection);
	//QMetaObject::invokeMethod(ra,"testFunction",Qt::QueuedConnection);

	return a.exec();
}

\end{minted}

\end{document}